---
title: 'Overriding individual code chunk options'
date: '2022-06-20'
code-fold: false
engine: knitr
---

I've just moved this entire blog from `{blogdown}` to `{quarto}`. It was mostly simple but for one of the blogposts I really wanted a way to override individual code chunk options. Let's make a toy .Rmd document to demonstrate this:

````{verbatim}
---
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)
```

This code chunk has `eval=TRUE` which I want to override

```{r, eval=TRUE}
2 + 2
```

````

- The default behaviour for code chunks is set in the `setup` chunk as `eval=FALSE`

- There's an individual code chunk with `eval=TRUE`

- I want to **override** the individual code chunk behaviour.

I tweeted about this and [Garrick Aden-Buie replied](https://twitter.com/grrrck/status/1538925669756944384?s=20&t=2Wh8cPaoSIgCwhD7yxUlOg) linking me to the documentation for option hooks. I was really pleased to then figure out this solution:

````{verbatim}
---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_hooks$set(force_not_eval = function(value) {
  value$eval <- FALSE
  value
}
)
knitr::opts_chunk$set(force_not_eval = TRUE)
```


```{r, eval=TRUE}
2 + 2
```
````


