---
title: "How far can you live from?"
date: '2024-11-08'
draft: true
execute:
  freeze: true
  message: false
  warning: false
  echo: false
  eval: false
code-fold: false
engine: knitr
filters:
  - line-highlight
---

```{r}
#| include: false
library("cjhRutils")
```

While working on the motorway services dataset I realised that I had an organic opportunity to look at edit distance

```{r}
library("tidyverse")
library("cjhRutils")
library("ggwordcloud")

data_for_edit_distance <- read_csv(quarto_here("data-for-edit-distance.csv"))

data_for_edit_distance %>% 
  add_count(cut = cut_width(n, 5), name = "n_in_cut") %>% 
  ggplot(aes(label = value,
             size = n,
             colour = cut)) +
  geom_text_wordcloud() +
  scale_size_area(max_size = 25) +
  # scale_colour_brewer(palette  = "Dark2") +
  scale_colour_viridis_d() +
  theme_minimal()
```




## How far can you live from a service station?

For this section we're only interested in how the crow flies distances - not driving distances. So, let's find the further point in the UK mainlaind from a service station... which is the further point on the border! There's a [lovely answer on StackOverflow](https://stackoverflow.com/a/76754196/1659890) that allowed me to build up this function:

```{r}
border_mainland_points <- data_sf_gb_mainland %>%
  st_cast("POINT")
```


```{r}
data_sf_service_locs

make_further_linestring_to_border <- function(data){
  
  point_service_station <- data %>% 
  st_as_sfc() %>% 
  st_as_sf() %>% 
  st_transform(crs = st_crs(data_sf_gb_mainland))
  
  distances <- border_mainland_points |>
    st_distance(point_service_station)
  
  max_dist_linestring <- border_mainland_points |>
    filter(as.logical(distances == max(distances))) |>
    bind_rows(point_service_station) |>
    summarise() |>
    st_cast("LINESTRING")
  
  max_dist_linestring %>% 
    mutate(name = data$name, length = st_length(x), .before = 0)
  
}

data_distances_services_to_border <- data_sf_service_locs %>% 
  mutate(id = row_number()) %>% 
  # filter(is_ireland == FALSE) %>% 
  slice(1) %>% 
  pull(id) %>% 
  map_dfr(~ make_further_linestring_to_border(slice(data_sf_service_locs, .x)))
```

```{r}
ggplot() +
  geom_sf(data = data_sf_simpler_mainland) +
  geom_sf(data = data_distances_services_to_border)
```


# !! THINGS TO DO !!

1. NOIN FOOD

```{r}
vec_not_food_shop <- c("BP", "Phone Bitz", "FoneBitz", "Tourist Information",
                       "Children's Play Area", "Game Arcade", "Regus Business Lounge", 
                       "Showers", "Lucky Coin", "Barber Shop", "Retail Shop",
                       "Westmorland shop")

```


# Dragons

Rules to clean

```{r}
data_raw_services %>% 
  mutate(value = case_when(
    property == "Picnic Area" ~ if_else(value == "yes", "TRUE", NA),
    .default = value
  )) %>% 
  filter(property == "Picnic Area")
```




