---
title: '#rmdnews updates 2019-03-06'
author: Charlie Joey Hadley
date: '2019-03-06'
slug: rmdnews-updates-2019-03-06
draft: yes
categories:
  - R
tags:
  - rmdnews
editor_options: 
  chunk_output_type: console
---

```{r emoji, include=FALSE}
library("emo")
```


It's hard keeping up to date with the awesome new things that we can do with [#rstats](https://twitter.com/search?q=%23rstats). There are new packages and interesting new use cases of existing `r emo::ji("package")` coming out all of the time. Mara Averick ([\@dataandme](https://twitter.com/dataandme)) is an incredible source of overviews and introductions to awesome stuff, if you don't already follow her then do so immediately.

But how about `r emo::ji("package")` updates? We often upgrade a package in a rush when we're told the `r emo::ji("package")` we want to install requires a new version of X. This means we miss out on potentially useful (or exciting) new features of packages that we already know and love.

So, I suggest as a community we use [#rmdnews](https://twitter.com/search?q=%23rstats) to tag tweets and blogposts that are solely to do with updates to existing packages.

Why this hashtag? Most^[citation] R packages contain a NEWS.md file which is where package developers are encouraged ([but **not** required](http://r-pkgs.had.co.nz/release.html#news)) to document updates and new features in their `r emo::ji("package")`. You find the NEWS.md file in one of two places:

- On CRAN in the DESCRIPTION file, e.g. [https://cran.r-project.org/web/packages/dplyr/index.html](https://cran.r-project.org/web/packages/dplyr/index.html)

- If the package lives on Github, it's extremely likely to have a NEWS.md file in the repo, e.g. https://github.com/tidyverse/dplyr/blob/master/NEWS.md 

# #rmdnews 2019-03-06

To get into the spirit of things, here are the new things I noticed in packages I needed to update to install tidyjson.

## forcats v0.4.0 #rmdnews

I love [`forcats`](https://forcats.tidyverse.org/). It massively simplifies manipulating factors, particularly within a tidyverse workflow.

### New functions

`fct_cross()` is an entirely new function that allows us to combine together multiple factors. It's similar to `purrr::cross` which creates all combinations of list elements. The functionality was [suggested back in July 2018](https://github.com/tidyverse/forcats/issues/136) along with this great example use case:

```{r}
library("tidyverse")
gss_cat %>% 
  select(year, marital, race, rincome) %>%
	mutate(
		marital_race = fct_cross(marital, race, sep = " & ")
	)
```

`fct_match` is an entirely new function that allows us to test if specific levels exist inside a factor. This allows us to filter by multiple conditions:

```{r}
gss_cat %>%
  filter(fct_match(marital, c("Married", "Divorced")))
```

This functionality was [suggested back in April 2018](https://github.com/tidyverse/forcats/issues/126) so as to explicitly provide errors if levels do not exist in the factor. Compare the output using `fct_match` and `%in%` below:

```{r, eval=FALSE}
gss_cat %>%
  filter(fct_match(marital, c("Married", "Divorceeeed")))
```

```{r, eval = FALSE}
gss_cat %>%
  filter(marital %in% c("Married", "Divorceeeed"))
```

### Improvements

[^citation]: Citation needed. Or better yet, data.
